Random number generation and applications on GPU’s in R?

### 2021-10-08 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
paper 1 update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2896 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

myStreamsCpu <- createStreamsCpu(n=4)

@@

myStreamsGpu2 = createStreamsGpu(n=4)

@@

Now we use the streams created in section \ref{create_streams} to generate a vector \code{sim_1} of 6 i.i.d. $U (0,1)$ random numbers using the function \fct{runif}. If \code{streams} and \code{Nglobal} are both missing, then by default \code{Nglobal} is set as (64,8), and a `vclMatrix' of \code{prod(Nglobal)} streams is supplied. The output is an S4 object (a `\code{vclVector}' or a `\code{vclMatrix}'), to view the generated random numbers, we need to convert it to an R vector or matrix, in this way, the random numbers are moved from the global memory of device to the host.
sim_1 <- runif(n = 6, streams = myStreamsGpu1, Nglobal = c(1, nrow(myStreamsGpu1)),
    type = "double")
as.vector(sim_1)
## [1] 0.7353245 0.5180770 0.6142074 0.2319392 0.1100781 0.3619766
The arguments of the \fct{gpuRandom::runif} are described as follows:
  \item \code{n}: a vector of length 2 specifying the row and column number if to create a matrix, or a number specifying the length if to create a vector.
  \item \code{streams}: streams for random number generation.
  \item \code{Nglobal}: global index space. Default is set as $(64,8)$.
  \item \code{type}: ``double'' or ``float'' format of generated random numbers.
Next we show that by using \code{myStreamsGpu2} which is the same as \code{myStreamsGpu1},  \fct{runif} returns the same sequence of random numbers as \code{sim_1}. 
sim_2 = runif(n = 7, streams = myStreamsGpu2, Nglobal = c(1, nrow(myStreamsGpu1)),
    type = "double")
as.vector(sim_2)
## [1] 0.7353245 0.5180770 0.6142074 0.2319392 0.1100781 0.3619766
## [7] 0.6487742
Reusing \code{myStreamsGpu1} will produce a vector different from \code{sim_1}. Because each time a random number is generated, the current state of the stream is changed (advances by one position), below shows the \code{myStreamsGpu1} after it being used.
t(matrix(as.matrix(myStreamsGpu1), nrow(myStreamsCpu), ncol(myStreamsCpu),
    dimnames = dimnames(myStreamsCpu)))
## current.g1.1  878672095 2113333390  502033783  739421137
## current.g1.2  240667857  559530223 1322587635 1475938232
## current.g1.3  240667857 1309565828 1964121530  730262207
## current.g2.1  642281259 1335994581 1949818481 1630192198
## current.g2.2 1069151070   61444481 1607232546  324551134
## current.g2.3  809054265  197003928 1462898381  795289868
Unlike the objects created by \proglang{R}, by default they remain in the memory after a restart. Streams created on GPU does not remain in the memory when a new \proglang{R} session begins. However, we can save the streams on the CPU in a file using the \fct{save} function, and later recall the streams with the \fct{load} function, in this way we can reproduce the same sequence of random numbers in simulations. Below is a trivial example, in which we demonstrate saving streams to a data file called \code{streams_from_GPU.rds} on CPU and then load it back and transfer it to GPU.
saveRDS(as.matrix(createStreamsGpu(n=4)), "myStreams.rds")
# Load the streams object as streams_saved
streams_saved <- vclMatrix(readRDS("myStreams.rds"))

@@

month_gpu<-vclMatrix(month,type="integer")
result_month <- fisher.sim(month_gpu, 1e6, streams=streams,
                   type="double", returnStatistics=TRUE,  Nglobal = c(256,128))
result_month$simNum
## [1] 1015808
result_month$counts
## [1] 410644
result_month$p.value
## [1] 0.4042541
\end{verbatim}
\end{kframe}
\end{knitrout}
We requested one million simulations on GPU while the actual number of simulation is 1015808, and got 403274 cases whose test statitsic is below the observed threshold, so the p-value from \fct{gpuRandom::fisher.sim} for this table is about 0.40375, which is close enough to the p-value from \fct{stats::fisher.test}: 0.40380.  
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
#using GPU
system.time(fisher.sim(month_gpu, 1e6, streams=streams, 
            type="double", returnStatistics=FALSE, Nglobal = c(256,128)))
##    user  system elapsed 
##   0.844   0.004   0.847
\end{verbatim}
\end{kframe}
\end{knitrout}
\fct{gpuRandom::fisher.sim} takes about 0.32 seconds, \pkg{gpuRandom} accounts for 2.2\% of the elapsed time on CPU.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
#using CPU
stats::fisher.test(month,simulate.p.value = TRUE,B=1015808)$p.value
## [1] 0.4037993
system.time(stats::fisher.test(month,simulate.p.value = TRUE,B=1015808))
##    user  system elapsed 
##  14.962   0.017  14.970
\end{verbatim}
\end{kframe}
\end{knitrout}
\fct{stats::fisher.test} takes about 9.8 seconds (time varies a little bit from executions) for 1015808 simulations, has a p-value of 0.4035.

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
observed_w= -gpuRandom::logfactSum(week, c(16,16))/almost.1
week_GPU<-gpuR::vclMatrix(week,type="integer")
result_weekgpu<-gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
                type="double",returnStatistics=TRUE,Nglobal = c(256,128))
result_weekgpu$simNum
## [1] 10027008
result_weekgpu$counts
## [1] 1242
result_weekgpu$p.value
## [1] 0.0001239652
\end{verbatim}
\end{kframe}
\end{knitrout}
With about ten million simulations, we get 1223 cases and a p-value around 0.0001224705. We summarized the results in Table \ref{tab:summary}.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
#using GPU
system.time(gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
            type="double", returnStatistics=FALSE,Nglobal = c(256,128)))
##    user  system elapsed 
##   4.555   0.012   4.564
\end{verbatim}
\end{kframe}
\end{knitrout}
\fct{stats::fisher.test} takes 90.1 seconds, while \fct{fisher.sim} takes about 1.90 seconds, the elapsed time is decreased to about 20\% of the time taken on CPU.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
#using CPU
fisher.test(week,simulate.p.value = TRUE,B=10010624)$p.value
## [1] 0.0001263657
system.time(fisher.test(week,simulate.p.value = TRUE,B=10010624))
##    user  system elapsed 
##  90.996   0.170  91.109
\end{verbatim}
\end{kframe}
\end{knitrout}
The ``week'' table has a much smaller p-value: around 0.0001208, which should require a larger number of simulations to get a more accurate p-value.

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
myParamsBatch
##      shape range variance nugget anisoRatio anisoAngleRadians
## [1,]  1.25  0.50      1.5      0          1         0.0000000
## [2,]  2.15  0.25      2.0      0          4         0.4487990
## [3,]  0.55  1.50      2.0      0          4         0.4487990
## [4,]  2.15  0.50      2.0      0          4        -0.4487990
## [5,]  2.15  0.50      2.0      0          2         0.7853982
paramsGpu = gpuRandom::maternGpuParam(myParamsBatch, type=theType)
\end{verbatim}
\end{kframe}
\end{knitrout}
Step 3, compute Mat\'ern covariance matrices using \fct{gpuRandom::maternBatch}, the returned Mat\'ern covariance matrices %$\Sigma_1, \Sigma_2, \dots, \Sigma_5$ correspond to the 5 parameter sets respectively, the matrices 
are each of size $4800 \times 4800$ and are stacked by row in the output \code{maternCov}. 
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
maternCov = vclMatrix(0, nrow(paramsGpu)*nrow(coordsGpu), 
                         nrow(coordsGpu),type=theType)
dim(maternCov)
## [1] 24000  4800
gpuRandom::maternBatch(maternCov, coordsGpu, paramsGpu,
                       Nglobal=c(128,64), Nlocal=c(16,4))
\end{verbatim}
\end{kframe}
\end{knitrout}
Step 4, the first argument \code{maternCov} in \fct{gpuRandom::cholBatch} specifies the matrix object to take Cholesky decomposition. Unit lower triangular matrices $L_i$'s are returned and stacked by row in \code{maternCov}, \code{diagMat} stores the diagonal values of each $D_i$ in a row, for example, if each batch $\Sigma_i$ is of size $n \times n$, then each batch $L_i$ is $n \times n$, and each batch $D_i$ is $1 \times n$.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
diagMat = vclMatrix(0, nrow(paramsGpu), ncol(maternCov), type = theType)
gpuRandom::cholBatch(maternCov, diagMat, numbatchD=nrow(myParamsBatch),
                     Nglobal= c(512, 8),
                     Nlocal= c(128, 8),
                     NlocalCache=2000)
\end{verbatim}
\end{kframe}
\end{knitrout}
% \begin{gather}
% \begin{bmatrix} \Sigma_{1} \\ \Sigma_2 \\ \Sigma_3 \\ \vdots
% \end{bmatrix}
%  \rightarrow
%  \begin{bmatrix}
%   L_{1} \\ L_{2} \\L_{3} \\ \vdots
%   \end{bmatrix} \text{and}
%   \begin{bmatrix}
%   D_{1} \\ D_{2} \\D_{3} \\ \vdots
%   \end{bmatrix}
% \end{gather}
Step 5, generate 2 standard Gaussian random vectors \code{zmatGpu}$=(Z_1, Z_2)$ using \fct{gpuRandom::rnorm}, in which \code{c(nrow(materCov),2)} specifies the number of rows and columns of \code{zmatGpu}.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
streamsGpu <- createStreamsGpu(n=128*64)
\end{verbatim}
\end{kframe}
\end{knitrout}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
zmatGpu = rnorm(c(nrow(maternCov),2), streams=streamsGpu, Nglobal=c(128,64),
                 type = theType)
\end{verbatim}
\end{kframe}
\end{knitrout}
Step 6, use \fct{gpuRandom::multiplyLowerDiagonalBatch} to compute $U = L * D^{(1/2)}* Z$ in batches, see the following illustration. \code{simMat} is the output matrix for $U$, \code{maternCov}, \code{diagMat}, and \code{zmatGpu} correspond to the matrices of $L$, $D$ and $Z=(Z_1, Z_2)$ respectively.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
simMat = vclMatrix(0, nrow(zmatGpu), ncol(zmatGpu), 
                   type = theType)

gpuRandom::multiplyLowerDiagonalBatch(
  simMat, maternCov, diagMat, zmatGpu,
  diagIsOne = 1L, # diagonal of L is one
  transformD = "sqrt", # take the square root of each element of D
  Nglobal,
  Nlocal,
  NlocalCache)
\end{verbatim}
\end{kframe}
\end{knitrout}
\begin{gather}
 \begin{bmatrix}  L_{1} \\ L_{2} \\L_{3} \\ \vdots
 \end{bmatrix} 
 *
  \begin{bmatrix}
   D_{1} \\ D_{2} \\D_{3} \\ \vdots
   \end{bmatrix} 
   *
   \begin{bmatrix}
   Z_{11} & Z_{12}
   \end{bmatrix}
  =
 \begin{bmatrix}
   L_1D_1Z_{11} & L_1D_1Z_{12} \\
   L_2D_2Z_{11} & L_2D_2Z_{12} \\
   L_3D_3Z_{11} & L_3D_3Z_{12} 
  \end{bmatrix}
\end{gather}
Step 7, Finally, plot the 10 simulated Gaussian random surfaces.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
library(raster)
simRaster = brick(myRaster, nl = ncol(simMat)*nrow(paramsGpu))
values(simRaster) = as.vector(as.matrix(simMat))
names(simRaster) = apply(expand.grid('par',1:nrow(paramsGpu), 
                                     'sim', 1:ncol(simMat)), 1, paste, collapse='')
\end{verbatim}
\end{kframe}
\end{knitrout}
\begin{figure}[H]
\centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
par(mar=rep(0.1, 4))
for(D in names(simRaster)) {
  plot(extent(simRaster))
  plot(simRaster[[D]], legend=FALSE, add=TRUE)
}
\end{verbatim}
\end{kframe}
\includegraphics[width=0.45\textwidth]{figure/maternplot-1} 
\includegraphics[width=0.45\textwidth]{figure/maternplot-2} 
\includegraphics[width=0.45\textwidth]{figure/maternplot-3} 
\includegraphics[width=0.45\textwidth]{figure/maternplot-4} 
\includegraphics[width=0.45\textwidth]{figure/maternplot-5} 
\includegraphics[width=0.45\textwidth]{figure/maternplot-6} 
\includegraphics[width=0.45\textwidth]{figure/maternplot-7} 
\includegraphics[width=0.45\textwidth]{figure/maternplot-8} 
\includegraphics[width=0.45\textwidth]{figure/maternplot-9} 
\includegraphics[width=0.45\textwidth]{figure/maternplot-10} 
\end{knitrout}
\caption{\label{fig:5} }
\end{figure}

>>>>

myStreamsCpu <- createStreamsCpu(n=4, seed=12345)

@@

myStreamsGpu2 = createStreamsGpu(n=4, seed=12345)

@@

Now we use the streams created in section \ref{create_streams} to generate two vectors of 6 i.i.d. $U (0,1)$ random numbers using the function \fct{runif}. If \code{streams} and \code{Nglobal} are both missing, then by default \code{Nglobal} is set as (64,8), and a `vclMatrix' of \code{prod(Nglobal)} streams is supplied. The output is an S4 object (a `\code{vclVector}' or a `\code{vclMatrix}'), to view the generated random numbers, we need to convert it to an R vector or matrix, in this way, the random numbers are moved from the global memory of device to the host. By using \code{myStreamsGpu2} which is the same as \code{myStreamsGpu1},  \fct{runif} returns the same sequence of random numbers. 
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
as.vector(gpuRandom::runif(n = 4, streams = myStreamsGpu1, Nglobal = c(1,
    nrow(myStreamsGpu1)), type = "double"))
## [1] 0.7353245 0.5180770 0.6142074 0.2319392
as.vector(gpuRandom::runif(n = 10, streams = myStreamsGpu2, Nglobal = c(1,
    nrow(myStreamsGpu1)), type = "double"))[-(1:6)]
## [1] 0.6487742 0.1112075 0.3661944 0.5018562
\end{verbatim}
\end{kframe}
\end{knitrout}
The arguments of the \fct{gpuRandom::runif} are described as follows:
\begin{itemize}
  \item \code{n}: a vector of length 2 specifying the row and column number if to create a matrix, or a number specifying the length if to create a vector.
  \item \code{streams}: streams for random number generation.
  \item \code{Nglobal}: global index space. Default is set as $(64,8)$.
  \item \code{type}: ``double'' or ``float'' format of generated random numbers.
\end{itemize}

Reusing \code{myStreamsGpu1} will produce a vector different from \code{sim_1}. Because each time a random number is generated, the current state of the stream is changed (advances by one position), below shows the \code{myStreamsGpu1} after it being used.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
t(matrix(as.matrix(myStreamsGpu1), nrow(myStreamsCpu), ncol(myStreamsCpu),
    dimnames = dimnames(myStreamsCpu)))
##                    [,1]       [,2]       [,3]       [,4]
## current.g1.1  240667857  559530223  502033783  739421137
## current.g1.2  240667857 1309565828 1322587635 1475938232
## current.g1.3      12345  336690377 1964121530  730262207
## current.g2.1 1069151070   61444481 1949818481 1630192198
## current.g2.2  809054265  197003928 1607232546  324551134
## current.g2.3      12345   85196284 1462898381  795289868
## initial.g1.1      12345  336690377  502033783  739421137
## initial.g1.2      12345  597094797 1322587635 1475938232
## initial.g1.3      12345 1245771585 1964121530  730262207
## initial.g2.1      12345   85196284 1949818481 1630192198
## initial.g2.2      12345  523477687 1607232546  324551134
## initial.g2.3      12345 2094976052 1462898381  795289868
\end{verbatim}
\end{kframe}
\end{knitrout}
Unlike the objects created by \proglang{R}, by default they remain in the memory after a restart. Streams created on GPU does not remain in the memory when a new \proglang{R} session begins. However, we can save the streams on the CPU in a file using the \fct{save} function, and later recall the streams with the \fct{load} function, in this way we can reproduce the same sequence of random numbers in simulations. Below is a trivial example, in which we demonstrate saving streams to a data file called \code{streams_from_GPU.rds} on CPU and then load it back and transfer it to GPU.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
saveRDS(as.matrix(createStreamsGpu(n=4)), "myStreams.rds")
# Load the streams object as streams_saved
streams_saved <- vclMatrix(readRDS("myStreams.rds"))
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

month_gpu<-vclMatrix(month,type="integer")
result_month <- fisher.sim(month_gpu, 1e6, streams=streams,
                   type="double", returnStatistics=TRUE,  Nglobal = c(256,128))
result_month$simNum
## [1] 1015808
result_month$counts
## [1] 410644
result_month$p.value
## [1] 0.4042541
\end{verbatim}
\end{kframe}
\end{knitrout}
We requested one million simulations on GPU while the actual number of simulation is 1015808, and got 403274 cases whose test statitsic is below the observed threshold, so the p-value from \fct{gpuRandom::fisher.sim} for this table is about 0.40375, which is close enough to the p-value from \fct{stats::fisher.test}: 0.40380.  
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
#using GPU
system.time(fisher.sim(month_gpu, 1e6, streams=streams, 
            type="double", returnStatistics=FALSE, Nglobal = c(256,128)))
##    user  system elapsed 
##   0.764   0.000   0.764
\end{verbatim}
\end{kframe}
\end{knitrout}
\fct{gpuRandom::fisher.sim} takes about 0.32 seconds, \pkg{gpuRandom} accounts for 2.2\% of the elapsed time on CPU.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
#using CPU
stats::fisher.test(month,simulate.p.value = TRUE,B=1015808)$p.value
## [1] 0.4037993
system.time(stats::fisher.test(month,simulate.p.value = TRUE,B=1015808))
##    user  system elapsed 
##  14.747   0.004  14.743
\end{verbatim}
\end{kframe}
\end{knitrout}
\fct{stats::fisher.test} takes about 9.8 seconds (time varies a little bit from executions) for 1015808 simulations, has a p-value of 0.4035.

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
observed_w= -gpuRandom::logfactSum(week, c(16,16))/almost.1
week_GPU<-gpuR::vclMatrix(week,type="integer")
result_weekgpu<-gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
                type="double",returnStatistics=TRUE,Nglobal = c(256,128))
result_weekgpu$simNum
## [1] 10027008
result_weekgpu$counts
## [1] 1242
result_weekgpu$p.value
## [1] 0.0001239652
\end{verbatim}
\end{kframe}
\end{knitrout}
With about ten million simulations, we get 1223 cases and a p-value around 0.0001224705. We summarized the results in Table \ref{tab:summary}.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
#using GPU
system.time(gpuRandom::fisher.sim(week_GPU, 1e7, streams=streams,
            type="double", returnStatistics=FALSE,Nglobal = c(256,128)))
##    user  system elapsed 
##   4.343   0.012   4.352
\end{verbatim}
\end{kframe}
\end{knitrout}
\fct{stats::fisher.test} takes 90.1 seconds, while \fct{fisher.sim} takes about 1.90 seconds, the elapsed time is decreased to about 20\% of the time taken on CPU.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
#using CPU
fisher.test(week,simulate.p.value = TRUE,B=10010624)$p.value
## [1] 0.0001263657
system.time(fisher.test(week,simulate.p.value = TRUE,B=10010624))
##    user  system elapsed 
##  90.524   0.155  90.629
\end{verbatim}
\end{kframe}
\end{knitrout}
The ``week'' table has a much smaller p-value: around 0.0001208, which should require a larger number of simulations to get a more accurate p-value.

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
myParamsBatch
##      shape range variance nugget anisoRatio anisoAngleRadians
## [1,]  1.25  0.50      1.5      0          1         0.0000000
## [2,]  2.15  0.25      2.0      0          4         0.4487990
## [3,]  0.55  1.50      2.0      0          4         0.4487990
## [4,]  2.15  0.50      2.0      0          4        -0.4487990
## [5,]  2.15  0.50      2.0      0          2         0.7853982
paramsGpu = gpuRandom::maternGpuParam(myParamsBatch, type=theType)
\end{verbatim}
\end{kframe}
\end{knitrout}
Step 3, compute Mat\'ern covariance matrices using \fct{gpuRandom::maternBatch}, the returned Mat\'ern covariance matrices %$\Sigma_1, \Sigma_2, \dots, \Sigma_5$ correspond to the 5 parameter sets respectively, the matrices 
are each of size $4800 \times 4800$ and are stacked by row in the output \code{maternCov}. 
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
maternCov = vclMatrix(0, nrow(paramsGpu)*nrow(coordsGpu), 
                         nrow(coordsGpu),type=theType)
dim(maternCov)
## [1] 24000  4800
gpuRandom::maternBatch(maternCov, coordsGpu, paramsGpu,
                       Nglobal=c(128,64), Nlocal=c(16,4))
\end{verbatim}
\end{kframe}
\end{knitrout}
Step 4, the first argument \code{maternCov} in \fct{gpuRandom::cholBatch} specifies the matrix object to take Cholesky decomposition. Unit lower triangular matrices $L_i$'s are returned and stacked by row in \code{maternCov}, \code{diagMat} stores the diagonal values of each $D_i$ in a row, for example, if each batch $\Sigma_i$ is of size $n \times n$, then each batch $L_i$ is $n \times n$, and each batch $D_i$ is $1 \times n$.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
diagMat = vclMatrix(0, nrow(paramsGpu), ncol(maternCov), type = theType)
gpuRandom::cholBatch(maternCov, diagMat, numbatchD=nrow(myParamsBatch),
                     Nglobal= c(512, 8),
                     Nlocal= c(128, 8),
                     NlocalCache=2000)
\end{verbatim}
\end{kframe}
\end{knitrout}
% \begin{gather}
% \begin{bmatrix} \Sigma_{1} \\ \Sigma_2 \\ \Sigma_3 \\ \vdots
% \end{bmatrix}
%  \rightarrow
%  \begin{bmatrix}
%   L_{1} \\ L_{2} \\L_{3} \\ \vdots
%   \end{bmatrix} \text{and}
%   \begin{bmatrix}
%   D_{1} \\ D_{2} \\D_{3} \\ \vdots
%   \end{bmatrix}
% \end{gather}
Step 5, generate 2 standard Gaussian random vectors \code{zmatGpu}$=(Z_1, Z_2)$ using \fct{gpuRandom::rnorm}, in which \code{c(nrow(materCov),2)} specifies the number of rows and columns of \code{zmatGpu}.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
streamsGpu <- createStreamsGpu(n=128*64)
\end{verbatim}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in vectorToVCL(data, 4L, context\_index - 1): ViennaCL: FATAL ERROR: CL\_OUT\_OF\_RESOURCES \\\#\# \ ViennaCL tried to launch a compute kernel, but the device does not provide enough resources. Try changing the global and local work item sizes.\\\#\# If you think that this is a bug in ViennaCL, please report it at viennacl-support@lists.sourceforge.net and supply at least the following information:\\\#\# \ * Operating System\\\#\# \ * Which OpenCL implementation (AMD, NVIDIA, etc.)\\\#\# \ * ViennaCL version\\\#\# Many thanks in advance!}}\end{kframe}
\end{knitrout}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
zmatGpu = rnorm(c(nrow(maternCov),2), streams=streamsGpu, Nglobal=c(128,64),
                 type = theType)
\end{verbatim}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in nrow(streams): object 'streamsGpu' not found}}\end{kframe}
\end{knitrout}
Step 6, use \fct{gpuRandom::multiplyLowerDiagonalBatch} to compute $U = L * D^{(1/2)}* Z$ in batches, see the following illustration. \code{simMat} is the output matrix for $U$, \code{maternCov}, \code{diagMat}, and \code{zmatGpu} correspond to the matrices of $L$, $D$ and $Z=(Z_1, Z_2)$ respectively.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
simMat = vclMatrix(0, nrow(zmatGpu), ncol(zmatGpu), 
                   type = theType)
\end{verbatim}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in h(simpleError(msg, call)): error in evaluating the argument 'x' in selecting a method for function 'nrow': object 'zmatGpu' not found}}\begin{verbatim}
gpuRandom::multiplyLowerDiagonalBatch(
  simMat, maternCov, diagMat, zmatGpu,
  diagIsOne = 1L, # diagonal of L is one
  transformD = "sqrt", # take the square root of each element of D
  Nglobal,
  Nlocal,
  NlocalCache)
\end{verbatim}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in gpuRandom:::multiplyLowerDiagonalBatchBackend(output, L, D, B, : object 'simMat' not found}}\end{kframe}
\end{knitrout}
\begin{gather}
 \begin{bmatrix}  L_{1} \\ L_{2} \\L_{3} \\ \vdots
 \end{bmatrix} 
 *
  \begin{bmatrix}
   D_{1} \\ D_{2} \\D_{3} \\ \vdots
   \end{bmatrix} 
   *
   \begin{bmatrix}
   Z_{11} & Z_{12}
   \end{bmatrix}
  =
 \begin{bmatrix}
   L_1D_1Z_{11} & L_1D_1Z_{12} \\
   L_2D_2Z_{11} & L_2D_2Z_{12} \\
   L_3D_3Z_{11} & L_3D_3Z_{12} 
  \end{bmatrix}
\end{gather}
Step 7, Finally, plot the 10 simulated Gaussian random surfaces.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
library(raster)
simRaster = brick(myRaster, nl = ncol(simMat)*nrow(paramsGpu))
\end{verbatim}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in h(simpleError(msg, call)): error in evaluating the argument 'x' in selecting a method for function 'ncol': object 'simMat' not found}}\begin{verbatim}
values(simRaster) = as.vector(as.matrix(simMat))
\end{verbatim}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in h(simpleError(msg, call)): error in evaluating the argument 'x' in selecting a method for function 'as.matrix': object 'simMat' not found}}\begin{verbatim}
names(simRaster) = apply(expand.grid('par',1:nrow(paramsGpu), 
                                     'sim', 1:ncol(simMat)), 1, paste, collapse='')
\end{verbatim}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in h(simpleError(msg, call)): error in evaluating the argument 'x' in selecting a method for function 'ncol': object 'simMat' not found}}\end{kframe}
\end{knitrout}
\begin{figure}[H]
\centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
par(mar=rep(0.1, 4))
for(D in names(simRaster)) {
  plot(extent(simRaster))
  plot(simRaster[[D]], legend=FALSE, add=TRUE)
}
\end{verbatim}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): object 'simRaster' not found}}\end{kframe}
\end{knitrout}
\caption{\label{fig:5} }
\end{figure}
